---
title: "hw04_Tidy-data-joins"
author: "Margot"
date: "08/10/2019"
output:
  html_document: default
  pdf_document: default
---
```{r, include=FALSE}
library(tidyverse)
library(gapminder)
library(DT)
```

## Exercise 1: Univariate Data Reshaping

### Option 1

1. Make a tibble with one row per year, and columns for life expectancy for two or more countries.
```{r}
gap_uni <- gapminder %>% filter(country=="Canada" |country== "China") %>% 
  select(country, year, lifeExp) %>% 
  pivot_wider(id_cols = year,
              names_from = country,
              values_from = lifeExp)
datatable(gap_uni)
```

2. Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.

```{r}
ggplot(gap_uni, aes(Canada, China))+
  geom_point()
```

3. Re-lengthen the data.

```{r}
gap_uni %>% pivot_longer(cols = c("Canada", "China"),
                             names_to = "country",
                             values_to = "lifeExp") %>% 
  datatable()
```

## Exercise 2: Multivariate Data Reshaping

### Option 1

1. Make a tibble with one row per year, and columns for life expectancy and GDP per capita (or two other numeric variables) for two or more countries.

```{r}
gap_multi <- gapminder %>% filter(country=="Cambodia" | country=="Egypt") %>% 
  select(year, country, lifeExp, gdpPercap) %>% 
  pivot_wider(id_cols = year,
              names_from = country,
              values_from = c(lifeExp, gdpPercap))
datatable(gap_multi)
```

2. Re-lengthen the data.

```{r}
gap_multi %>% pivot_longer(cols = -year,
                           names_to = c(".value", "country"),
                           names_sep = "_") %>% 
  datatable()
```

## Exercise 3: Table Joins
```{r, include=FALSE}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```

3.1 For each guest in the guestlist (guest tibble), add a column for email address, which can be found in the email tibble.
```{r}
emails <- email %>% separate_rows(guest, sep = ", ") %>% # seperate a collapsed column into multiple rows. Notice the sep is one comma and one white space
  mutate(name = guest) %>% 
  select(name, email)
emails
```
```{r}
guest %>% 
  left_join(emails, by="name") %>% 
  datatable()
```


3.2 Who do we have emails for, yet are not on the guestlist?
```{r}
emails %>% anti_join(guest, by="name")
```


3.3 Make a guestlist that includes everyone we have emails for (in addition to those on the original guestlist).

```{r}
guest %>% full_join(emails, by="name") %>% 
  datatable()
```

