---
title: "Assignment 02: Explore Gapminder and use dplyr"
output: github_document
---
<!--To run the code but show nothing-->
```{r include=FALSE}
library(gapminder)
library(tidyverse)
library(DT)
```

## Exercise 1: Basic __dplyr__

### 1.1 
Use __filter()__ to subset the __gapminder__ data to three countries of your choice in the 1970’s.
```{r}
gapminder %>% 
  filter(country == "Canada" | country == "Chile" | country == "Cambodia",
         year < 1980 & year > 1969)
```


### 1.2 

Use the pipe operator __%>%__ to select “country” and “gdpPercap” from your filtered dataset in 1.1.
```{r}
gapminder %>% 
  filter(country == "Canada" | country == "Chile" | country == "Cambodia",
         year < 1980 & year > 1969) %>% 
  select(country, gdpPercap)
```

### 1.3 

Filter gapminder to all entries that have experienced a drop in life expectancy. Be sure to include a new variable that’s the increase in life expectancy in your tibble. Hint: you might find the __lag()__ or __diff()__ functions useful.

```{r}
gapminder %>% 
  arrange(country, year) %>%
  mutate(lifeIncrease = lifeExp - lag(lifeExp)) %>% # lag(lifeExp) finds the previous values of lifeExp
  filter(lifeIncrease < 0, year != 1952)#exclude the first entry of each country, where the lifeIncrease is the difference between two countries 
```

### 1.4 

Filter gapminder so that it shows the max GDP per capita experienced by each country. Hint: you might find the __max()__ function useful here.
```{r}
gapminder %>% 
  group_by(country) %>% 
  mutate(maxgpdPercap = max(gdpPercap)) %>% 
  filter(gdpPercap == maxgpdPercap)
```

### 1.5

Produce a scatterplot of Canada’s life expectancy vs. GDP per capita using ggplot2, without defining a new variable. That is, after filtering the gapminder data set, pipe it directly into the __ggplot()__ function. Ensure GDP per capita is on a log scale.

```{r}
gapminder %>% 
  filter(country == "Canada") %>% 
  ggplot(aes(lifeExp, log(gdpPercap))) +
  geom_point()
```


## Exercise 2: Explore individual variables with __dplyr__

Pick one categorical variable and one quantitative variable to explore. Answer the following questions in whichever way you think is appropriate, using dplyr:

- What are possible values (or range, whichever is appropriate) of each variable?

- What values are typical? What’s the spread? What’s the distribution? Etc., tailored to the variable at hand.

I picked one categorical variable __country__ and one quantatative variable __pop__.

### country

The possible values of __country__:
```{r}
gapminder %>% 
  distinct(country)
```

The count of each possible value of __country__
```{r}
gapminder %>% 
  group_by(country) %>% 
  count()
```

### pop

The range of __pop__:
```{r}
range(gapminder$pop)
```

The spread of __pop__:
```{r}
summary(gapminder$pop)
```

The distribution of __pop__:
```{r}
ggplot(gapminder, aes(pop/1000000))+
  geom_histogram(binwidth = 50)
```

The distribution of __pop__ is very right skewed. Most values are below 150 million, while a few outliers are larger than 1 billion.

## Exercise 3: Explore various plot types

- A scatterplot of two quantitative variables.

- One other plot besides a scatterplot.

### Scatterplot
For this exercise, I used __mtcars__. I made a scatterplot of __mpg__(Miles/(US) gallon) and __wt__(Weight (1000 lbs)).
The plot shows a negative relationship between the fuel economy and the weight.
```{r}
cyl_number <- as.character(mtcars$cyl) #Change the type of cyl into character because there are only three discrete values.
ggplot(mtcars, aes(mpg, wt))+
  geom_point(aes(color= cyl_number))
```

### Boxplot
I made a boxplot showing the distribution of fuel economy(mpg) of cars with different number of cylinders.
```{r}
ggplot(mtcars, aes(cyl_number,mpg))+
  geom_boxplot()
```

