---
title: "HW05 Factor and figure management"
author: "Margot"
date: "17/10/2019"
output: 
  html_document:
    theme: paper
    toc: true
    keep_md: true
---
```{r, include=FALSE}
library(here)
library(gapminder)
library(tidyverse)
library(forcats)
library(knitr)
```

## Exercise 1: Explain the value of the here::here package
- `here::here` makes it easier and safer to write the file paths in different operating systems.
- It also makes it easier to manage sub-directories in a project. We can get data from other sub-directories and save output to other sub-directories.
- It allows people to run your code on their machine and supports reproducibility.

## Exercise 2: Factor management

### 2.1 Drop factor / levels

```{r}
class(gapminder$continent)
class(gapminder$country)
```
In `gapminder`, `country` and `continent` are factors.

```{r}
four_continent <- gapminder %>% filter(continent != "Oceania")
nlevels(four_continent$continent)
```

```{r}
continent_dropped <- four_continent %>% droplevels()
nlevels(continent_dropped$continent)
```

There are 5 levels of `continent`. After applying `droplevels()`, there are 4 left.

### 2.2 Reorder levels based on knowledge from data.

```{r}
df1 <-gapminder %>% 
  filter(country %in% c("Afghanistan", "Cambodia",  "India", "Japan", "Mexico")) 
df1 %>% arrange(country) %>% DT::datatable()
```

```{r}
df1 %>% ggplot()+
  geom_boxplot(aes(country, lifeExp))+
  theme_bw()
```

In the original tibble, countries are ordered alphabetically.

```{r}
df2 <- df1 %>% arrange(fct_reorder(country, lifeExp, max, .desc =TRUE)) 
DT::datatable(df2)
```

```{r}
df2 %>% ggplot()+
  geom_boxplot(aes(fct_reorder(country, lifeExp, max, .desc =TRUE), lifeExp))+
  xlab("country")+
  theme_bw()
```

In the reordered tibble, countries are ordered by the maximum of `lifeExp`, from the biggest to the smallest.

## Exercise 3: File input/output (I/O)

I use `df1`, a tibble I created in the last section for Exercise 3. First, I export it to disk.

```{r}
write_csv(df1, here::here("hw05", "df1.csv"))
```

Then reload `df1` back.

```{r}
my_data <- read_csv(here::here("hw05", "df1.csv"))
```

```{r}
my_data %>% 
  arrange(fct_relevel(country, "Mexico")) %>% 
  DT::datatable()
```

After applying `relevel()`, Mexico became the first country in the table.

## Exercise 4: Visualization design

The original vis:

```{r message=FALSE, warning=FALSE}
p <- gapminder %>% # store image in p for storing in the exercise
  group_by(year) %>% 
  mutate(life_mean = mean(lifeExp)) %>% 
  ggplot(aes(lifeExp))+
  facet_wrap(~ continent)+
  geom_histogram(aes(fill=lifeExp>life_mean))+
  scale_fill_discrete("", labels=c("Lower than average lifeExp", "Higher than average lifeExp"))
p
```

The new vis:

```{r message=FALSE, warning=FALSE}
gapminder %>% 
  group_by(year) %>% 
  mutate(life_mean = mean(lifeExp)) %>% 
  mutate(continent=fct_reorder(continent, lifeExp, median)) %>% #order continents by median lifeExp
  ggplot(aes(lifeExp))+
  facet_wrap(~ continent)+
  geom_histogram(aes(fill=lifeExp>life_mean))+
  scale_fill_discrete("Legend", labels=c("Lower than average lifeExp", "Higher than average lifeExp")) +
  ylab("Count of records of lower/higher life expectancy")+
  xlab("Life expectancy")+
  labs(title = "Life Expectancy in Different Continent")+
  theme_bw()
```

### Things I did differently:
- Add label, title, and legend to make the figure more informative. 
- Remove background to reduce distraction. 
- Reordered the continents to make the pattern more obvious.

## Exercise 5: Writing figures to file

I don't want to save the last image. So I use `plot=` to point out the image I want do save.

```{r message=FALSE, warning=FALSE}
ggsave(here("hw05", "life_exp.png"), plot=p, width = 7, height = 4)
```

Here is the [*Old Figure*](life_exp.png)
